Symmetric Circuits for Rank Gates

### 2018-03-30 sections/transparency-and-unique-extensions.tex
Major Updates to Section 5 (and a complete check of the first part of this section). Fixed a few typos in Section 3.

In the statement of the main theorem of this paper, rather than working with
general symmetric rank-circuits, we instead restrict ourselves to families of
\emph{transparent} circuits. In this section we justify this restriction,
showing first that the a great number of very useful circuit properties (e.g.
syntactic-equivalence) are polynomial-time decidable for transparent circuits
and, second, that deciding most of these circuit properties for general circuits
(with non-symmetric gates) is at least as hard as the graph-isomorphism
problem. Since the proof of the main theorem makes use of the polynomial-time
decidability of many of these properties, these results together justify the
inclusion of the transparency restriction.

\subsection{The Usefulness of Transparency}
In the this subsection we show that a number of important circuit properties are
polynomial-time decidable for transparent circuits (e.g. symmetry, the
syntactic-equivalence relation, etc.). We also provide polynomial-time
algorithms for computing the action of a permutation on a symmetric circuit, as
well as for computing the orbits and supports of gates, and elements of
universes of gates, in a circuit.

We use the polynomial-time decidability of many of these properties throughout
this paper, and particularly for our translation of circuits into formulas. For
example, the polynomial-time decidability of the syntactic-equivalence relation
enables us to transform transparent circuits into circuits with unique labels
(and so unique extensions) in polynomial-time. This result allows us, when
working with $\PT$-uniform families of transparent circuits, to assume without a
loss of generality that these circuits have unique labels and unique extensions.
This is a crucial step in the proof of our main result, as the unique labels
requirement is needed to apply the support theorem, and the unique extensions
property is necessary for applying (and even defining) a great many technical
tools (e.g. supports and supporting partitions) used extensively throughout this
paper.

@@

\begin{lem}
  There is an algorithm that takes as input a transparent circuit $C$ and
  outputs the syntactic-equivalence relation on the gates of $C$. The algorithm
  runs in time polynomial in the size of $C$.
  \label{lem:transparent-syntactic-equiv}
\end{lem}
\begin{proof}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a transparent
  $(\mathbb{B}, \rho)$-circuit of order $n$. We define the syntactic-equivalence
  relation on $G$ by induction on the depth of a gate We note that this
  definition by induction can be implemented as a construction using a dynamic
  program.

@@

  The above algorithm can be implemented by a dynamic program with the number of
  iterations bounded by the number of pairs of gates in $C$, i.e. $\vert C
  \vert$ choose $2$. It remains to show that each iterative step can be
  implemented so as to run in time polynomial in the size of $C$. The only
  condition that is not obviously polynomial-time decidable is checking if
  $L(g)$ is isomorphism-equivalent to $L(h)$ up to $\equiv$. We sketch an
  algorithm that decides this condition in time polynomial in the size of the
  circuit.

  Let $g$ and $h$ be two internal gates in $C$ and suppose we have defined the
  syntactic-equivalence relation for all gates of depth less than the depth of
  either $g$ or $h$. We test if for each gate in $g' \in H_g$, there exists
  $h_{g'} \in H_h$ such that $g' \equiv h_{g'}$ and $\vert \{g'' \in H_g : g''
  \equiv g'\} \vert = \vert \{h'' \in H_h : h'' \equiv h'\}$. If this test
  fails, we set $g \not\equiv h$.

  We then check if both $g$ and $h$ are symmetric gates or if both are
  non-symmetric gates. Suppose both $g$ and $h$ are symmetric gates. We have
  that $L(g)$ is isomorphism-equivalent to $L(h)$ up to $\equiv$ if, and only
  if, there exists a bijection $f : H_g \rightarrow H_h$ such that $f(g') \equiv
  g'$, for all $g' \in H_g$ if, and only if, for each gate in $g' \in H_g$,
  there exists $h' \in H_h$ such that $g' \equiv h'$ and $\vert \{g'' \in H_g :
  g'' \equiv g'\} \vert = \vert \{h'' \in H_h : h'' \equiv h'\}$. Since we have
  already checked this condition we set $g \equiv h$.

  Suppose $g$ and $h$ are non-symmetric gates. Then, since $C$ is transparent,
  both $g$ and $h$ have unique labels. We thus have that for each $g' \in H_g$
  there is exactly one $h_{g'} \in H_h$ such that $g' \equiv h_{g'}$, and so we
  can define the bijection $f : H_g \rightarrow H_h$ by $f(g') := h_{g'}$ for
  all $g' \in H_g$. It can be shown that $L(g)$ is isomorphism-equivalent to
  $L(h)$ up to $\equiv$ if, and only if, $f \circ L(g)$ is
  isomorphism-equivalent to $L(h)$ if, and only if, $L(h)^{-1} f L(g)$ acts on
  $\ind (g)$ like an element of $\aut(g)$. We have from
  Lemma~\ref{lem:aut-partial} that $\sigma := L(h)^{-1} f L(g)$ acts like an
  element of $\aut(g)$ if, and only if, for all $\vec{a}_R, \vec{b}_{R'} \in
  \ind(g)$, $\sigma$ acts like a partial automorphism on $(\vec{a}_R,
  \vec{b}_{R'})$. We thus conduct the test by iterating through every pair
  $(\vec{a}_R, \vec{b}_{R'}) \in \ind(g)$, and set $g \equiv h$ if, and only if,
  $\sigma$ acts like a partial automorphism on every such pair.
\end{proof}

It is sometimes useful to define one circuit from another circuit by
\emph{quotienting} a circuit by the syntactic equivalence relation. Intuitively,
a quotient circuit is defined by `merging' each syntactic-equivalence class into
a single gate and including a wire between two gates in the quotient circuit if,
and only if, there is a wire between an element of each equivalence class.

@@

We now show that taking the quotient of a circuit preserves important
properties, including the function computed by the circuit, whether the circuit
is symmetric, and has unique unique labels. We also show that the quotient of a
circuit is reduced, and hence has unique extensions.

@@

  Let $g \in G$. Suppose $[g]$ has unique labels. There exists $h \in [g]$ such
  that $L_\equiv([g]) = L(h) /_\equiv$. Since $L_\equiv([g])$ is injective,
  $L(h) /_\equiv$ must be injective and so $L(h)$ must be injective and no two
  child gates of $L(h)$ must be syntactically-equivalent. It follows that $h$
  has unique labels. Since $h \equiv g$ and $h$ has unique labels, it follows
  that $g$ has unique labels. Suppose $g$ has unique labels. Let $h \in [g]$ be
  such that $L_\equiv([g]) = L(h) /_\equiv$. Since $h \equiv g$ and $g$ has
  unique labels $h$ has unique labels and so $L_\equiv([g])$ has injective
  labels. Since each equivilence class in $C_\equiv$ is a singlton it follows
  that $[g]$ has unique labels.

  % If $[g]$ is an input gate then $\pi_\equiv [g] = [\pi g]$, and

  % Let $\mathcal{A}$ be a $\rho$-structure of size $n$ and let $\gamma$ be a
  % bijection from $U$ to $[n]$. We have from Lemma~\ref{} that for all $g, g'
  % \in
  % G$, if $g \equiv g'$ then $C[\gamma \mathcal{A}](g) = C[\gamma
  % \mathcal{A}](g')$. Moreover, we should note that for each internal gate

  % We show that for all $g \in G$, $C_\equiv [\gamma \mathcal{A}]([g]) =
  % C[\gamma
  % \mathcal{A}](g)$ by induction on the structure of the circuit. Suppose $g
  % \in
  % G$ is an input gate then, since every input gate is part of a singleton
  % equivalence class, we have that $C /_\equiv [\gamma \mathcal{A}]([g]) =
  % C[\gamma \mathcal{A}](g)$. Suppose $g \in G$ is an internal gate. Let $h \in
  % [g]$ such that $L /_\equiv ([g]) = L(h)/_\equiv$. Then $(L/_\equiv)^{\gamma
  % \mathcal{A}}([g])(x) = C /_\equiv[\gamma \mathcal{A}](L /_\equiv ([g])(x)) =
  % C [\gamma \mathcal{A}](L(h)(x)) = L^{\gamma \mathcal{A}}(h)(x) \equiv
  % L^{\gamma \mathcal{A}}(g) \circ \lambda (x)$ for some $\lambda \in \aut(g)$.
  % But then, since $\Sigma (g)$ is a structured function, we have that
  % $\Sigma(g)((L^{\gamma \mathcal{A}}(g) \lambda)) = \Sigma(g)(\L^{\gamma
  % \mathcal{A}})$. It follows that $C /_\equiv [\gamma\mathcal{A}]([g]) =
  % \Sigma /_\equiv ((L/_\equiv)^{\gamma mathcal{A}}([g])) = \Sigma (L^{\gamma
  % \mathcal{A}}(g)\lambda) = \Sigma (L^{\gamma \mathcal{A}}(g)) = C[\gamma
  % \mathcal{A}](g)$. It follows that $C/_\equiv [\gamma \mathcal{A}]
  % (\Omega/_\equiv (\vec{a})) = C [\gamma \mathcal{A}](\Omega (\vec{a}))$ for
  % all
  % $a \in [n]^q$. It follows that $C_{\equiv}$ and $C$ compute the same
  % function.

  % Suppose $C$ is symmetric. Suppose $C$ is symmetric. We have from For $\sigma
  % \in \sym_n$ let $\pi \in \aut(C)$ be an extension of $\sigma$. It can be
  % shown
  % that $\pi$ is invariant under $\equiv$, and hence we have that $\pi
  % /_{\equiv}$ is a function. Since each gate that is either an input or output
  % gate is a member of a singleton equivalence class we only need to check the
  % symmetry condition for internal gates. Suppose $g$ is an internal gate then
  % $\pi /_{\equiv} L /_{\equiv}([g]) (x) = [\pi L(g) (x)] = [L(\pi g) (\lambda
  % x)] = L([\pi g]) (\lambda x) = L(\pi /_{\equiv} [g]) (\lambda x)$ for some
  % $\lambda \in \aut(g)$. It follows that $\pi /_{\equiv}$ is an automorphism
  % of
  % $C_{\equiv}$ extending $\sigma$, and so $C_{\equiv}$ is symmetric.

  % Let $[g], [g'] \in G/_{\equiv}$. Suppose $[g] \equiv [g']$. If $g$ is an
  % input
  % or output gate we have that $g = g'$. Suppose $g$ is an internal gate. Then
  % there exists $h \in [g]$ and $h' \in [g']$ such that $L /_\equiv ([g]) =
  % L(h)
  % /_\equiv$ and $L /_\equiv ([g']) = L(h') /_\equiv$. We have that
  % $L_{\equiv}([g])$ is isomorphism-equivilent to $L_{\equiv}$

  % Let $g \in G$. Since every gate in $G /_{\equiv}$ is part of a singleton
  % equivalence class.

  % Then for each gate $g,
  % g' \in G$ such that $g \equiv g'$ we have that $C[\gamma\mathcal{A}](g) =
  % C[\gamma \mathcal{A}](g')$. If $g$ is an input gate we have that
  % $C[\gamma\mathcal{A}](g) = C /_\eqiv [\gamma \mathcal{A}]([g])$. Suppose
  % $[g] \in G /_{\equiv}$ is an internal gate. Let $L^{\gamma}([g]) : \ind(g)
  % \rightarrow \{0,1\}$ be defined by $L^{\gamma}([g]) (x) = C[\gamma
  % \mathcal{A}]([g])$
\end{proof}

It is worth noting that if $C$ is a transparent circuit we can compute a
quotient of $C$ in time polynomial in the size of $C$. Since the circuit is
transparent we can compute the syntactic-equivalence classes of $C$ in
polynomial time (see Lemma~\ref{lem:transparent-syntactic-equiv}). Picking a
representative of each equivalence class and wiring up the circuit in the
obvious way then yields the desired quotient circuit.

We now show that transparent circuits can be transformed in polynomial time into
equivalent circuit with unique labels, and hence unique extensions (see
Proposition~\ref{prop:unique-children-unique-extensions}), and that this
transformation preserves important properties of the circuit.

\begin{lem}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a transparent
  $(\mathbb{B}, \rho)$-circuit. There is an algorithm that that takes in such a
  circuit and outputs a $(\mathbb{B} \cup \mathbb{B}_{\std}, \rho)$-circuit $C'$
  such that $C$ and $C'$ compute the same function, $C'$ has unique labels, and
  if $C$ is symmetric then $C'$ is symmetric. Moreover, this algorithm runs in
  time polynomial in the size of the input circuit.
  \label{lem:transparent-unique}
\end{lem}
\begin{proof}
  Let $C_0 := \langle G_0, \Omega_0, \Sigma_0, \Lambda_0, L_0 \rangle$ be the
  quotient of $C$. If $C_0$ does not contain the constant gates $g_0$ and $g_1$
  such that $\Sigma(g_0) = 0$ and $\Sigma (g_1) = 1$ we construct a new circuit
  from $C_0$ by adding these gates. We abuse notation and also call this new
  circuit $C_0$. The addition of these constant gates to the circuit does not
  alter the function computed by the circuit, nor does it effect the symmetry of
  the circuit or weather it has unique labels.

@@

  We will define $C_1$ from $C_0$ by adding a designated gate that always
  evaluates to $1$ and which is part of a singleton syntactic-equivalence class,
  and then adding a wire from that gate to each $\AND$-gate in the circuit. This
  process ensures that each $\AND$-gates in $C_1$ has a fan-in of at least two.
  The construction of $C'$ from $C_1$ involves adding in a number of
  single-input $\AND$-gates, and so the construction of $C_1$ will allow us to
  distinguish these new gates from those in $C_1$.

  We now define the circuit $C_1$ formally. Let $C_1 := \langle G_1, \Omega_1,
  \Sigma_1 , \Lambda_1, L_1 \rangle$ be defined as follows. Let $G_1 = G_0
  \uplus \{ g_\lor\}$. Let $\Omega_1 = \Omega_0$ and $\Lambda_1 = \Lambda_0$.
  Let $g \in G_1$. If $g = g_\lor$ then $\Sigma_1 (g) = \OR[2]$, $L_1(g)(1) =
  g_0$, and $L_1(g)(2) = g_1$. If $g \in G_0$, $\Sigma(g) = \OR[2]$ and $H_g =
  \{g_0, g_1\}$ then $\Sigma_1 (g) = \OR[3]$, $L_1(g)(1) = L_0(g)(1)$,
  $L_1(g)(2) = L_0(g)(2)$, and $L_1(g)(3) = g_\lor$. If $g \in G_0$ and $g =
  \AND[k]$ for some $k \in \nats$ then $\Sigma_1(g) = \AND[k+1]$ and $L_1(g)(i)
  = L_0(g)(i)$ for all $i \in [k]$ and $L_1(g)(k+1) = g_\lor$. Otherwise let
  $\Sigma_1(g) = \Sigma_0(g)$ and $L_1(g) = L_0(g)$.

  More informally, we define $C_1$ from $C_0$ by adding in an $\OR$-gate
  $g_\lor$ that always evaluates to one, and then adding a wire from that gate
  to all $\AND$-gates in the circuit, and also a wire from $g_\lor$ to another
  $\OR$-gate in order to ensure that $g_\lor$ is part of a singleton
  syntactic-equivalence class in $C_1$. The important point to note is that each
  $\AND$-gate in $C_1$ has fan-in at least two, a property we will use to
  construct $C'$. We now show that the construction of $C_1$ preserves the
  relevant circuit properties.

@@

  Suppose $C$ is symmetric. From Lemma~\ref{lem:quotient-circuits-preserve} it
  follows that $C_0$ is symmetric. Let $\sigma \in sym_n$ and let $\pi_0 \in
  \aut(C_0)$ be an extension of $\sigma$. We define a bijection $\pi_1 : G_1
  \rightarrow G_1$ such that $\pi_1 (g) = \pi_0(g)$ for all $g \in G_0$ and
  $\pi_1 (g_\lor) = g_\lor$. It is easy to see that $\pi_1$ is an automorphism
  of $C_1$ extending $\sigma$. It follows that $C_1$ is symmetric.

  We have from Lemma~\ref{lem:quotient-circuits-preserve} that $C_0$ is reduced.
  Since $g_\lor$ is the only two input $\OR$-gate with exactly the two constant
  gates as children, it follows that $g_\lor$ is contained in a singleton
  syntactic-equivalence class. It can be shown by induction that if $g, g' \in
  G_0$ are syntactically equivalent in $C_1$ then they must be
  syntactically-equivalent in $C_0$. It follows that if $g$ and $g'$ are
  syntactically-equivalent in $C_1$ then $g = g'$. We thus have that $C_1$ is
  reduced.

>>>>

Many of the properties of circuits central to work of this paper, and
particularly those that reference circuit automorphisms or
syntactic-equivalence, make references to the existence of a isomorphism between
pairs of structures associated with the gates of the circuit. We recall that
each gate in a circuit is associated with a structure with the elements of the
relations of that structure labelled by the children of that gate. These
properties usually depend on the existence of an isomorphism that agrees with a
given permutation of the gates labelling the structure. This introduces a
potential difficulty as there is no know polynomial-time algorithm for the
general isomorphism problem, but we we require many of these circuit properties
to be polynomial-time decidable in order to define our translation from circuits
to formulas (see Section~\ref{sec:circuits-to-formulas}). This is not a problem
if we restrict our attention to circuits with symmetric gates, as Anderson and
Dawar~\cite{AndersonD17} do, as in this case the structures of interest are
sets, and set equivalence is clearly decidable in polynomial-time, but when we
allow for circuits to be defined over bases including non-symmetric functions it
is not obvious that these properties can be decided in polynomial-time.

This is the reason we introduced the notion of a transparent circuit. The idea
is that for transparent circuits the structures associated with each
non-symmetric gate have the property that each element of each relation is
labelled by a `unique' gate (i.e.\ such that no elements are labelled by gates
from the same syntactic-equivalence class). In this case a permutation of the
gates in the circuit defines exactly one function between the elements of the
relations of these structures, and so checking for the existence of an
isomorphism that agrees with the permutation is reduced to the problem of
determining if a given function is in fact an isomorphism. This problem is
polynomial-time tractable.

In the first part of this section we prove that the requisite circuit properties
are indeed polynomial-time decidable for transparent circuits. We then use these
results to define a polynomial-time computable translation from transparent
circuits to equivalent circuits with unique labels. As discussed in
Section~\ref{sec:symm-circ}, this translation allows us to restrict our
attention from transparent circuits to circuits with unique labels when working
with families of circuits without a loss of generality. This restriction is
important for two reasons. First, circuits with unique labels have unique
extensions, which enables us to apply the support theorem. Second, as we will
show in this section, all relevant symmetry-related circuit properties are
polynomial-time decidable for circuits with unique labels. We use all of these
properties in order to define our translation from circuits to formulas in
Section~\ref{sec:circuits-to-formulas}.

In the second part of this section we discuss more formally the relationship
between the problem of deciding many of these natural circuit properties and the
graph-isomorphism problem. In particular, we present reductions from the
graph-isomorphism problem to the problems of deciding each of these circuit
properties for general circuits. It follows that deciding most of these circuit
properties is hard as the graph isomorphism problem -- a problem not known to be
in $\PT$. We also show that most of these hardness results are quite robust, in
that even if we restrict our attention to other natural classes of circuits
(e.g.\ circuits with unique children or circuits with injective labels) these
decision problems remain at least as hard as the graph-isomorphism problem.
In this sense, while the first part of this section provides evidence in favour of the
sufficiency of the restriction to transparent circuits, the second part provides
evidence in favour of its necessity.

% In Section~\ref{sec:circuits-to-formulas} we use the polynomial-time
% decidability of these properties, along with the Immerman Vardi theorem, in
% order to define a translation from circuit families to formulas of $\FPR$. The
% results in this section thus justify the restriction to trans

% the results in this section are useful both are useful needed for proving the
% main result and help explain why it is that we need to restrict ourselves to
% transparent circuits, rather than working with general circuits.

% Since the proof of the main result makes use of the polynomial-time
% decidability of many of these properties, these results help explain both why
% transparent circuits suffice for the proof of our main result and why general
% circuits

% The restriction to transparent circuits is not found in the analogous circuit
% characterisation of $\FPR$ by Anderson and Dawar~\ref{AndersonD17}.

% In the first part of this section we show that a number of natural, and
% useful, circuit properties (e.g. symmetry, syntactic-equivalence) are
% polynomial-time decidable for transparent circuits. In
% Section~\ref{sec:circuits-to-formulas} we use the polynomial-time decidability
% of these properties, along with the Immerman-Vardi theorem, in order to define
% a translation from families of circuits to formulas. We show in
% Section~\ref{sec:formulas-to-circuits} that we can translate formulas to
% families of transparent circuits. It follows that $\PT$-uniform families of
% circuits suffice for our purposes.

% In the second part of this section we justify the necessity of this
% restriction, showing that most of these circuit properties for general
% circuits (with non-symmetric gates) are at least as hard as the
% graph-isomorphism problem. Since the proof of the main theorem makes use of
% the polynomial-time decidability of many of these properties, these results
% together justify the inclusion of the transparency restriction.

% The second part of this

% used to translate circuit families

% Importantly, our proof of the main result makes use of the polynomial-time
% decidability of these properties.

% many natural show that a great number of

% justify this restriction to transparent circuits, first showing that In this
% section we justify this restriction, showing first that the a great number of
% very useful circuit properties (e.g. syntactic-equivalence) are
% polynomial-time decidable for transparent circuits

% The transparency condition is a restriction on how non-symmetric gates are
% wired into the circuit, and so all circuits with symmetric gates are
% transparent. The circuit we have introduced includes

% We have thought of a gate $g$ with a vocabulary $\tau$ in a circuit $C$ as
% labelling its children using elements of the $\ind(g)$ (i.e\ the elements of
% the relations of the $\tau$-structure defined at $g$). We could invert this
% thinking, and instead think of the children of $g$ as labelling the elements
% of $ind(g)$, and so defining at $g$ a $\tau$-structure with a labelling on the
% elements of its relations. In particular, if $g$ is a rank-gate then $g$
% defines a complete bipartite graph labelled by the children of $g$.

% The idea is that a gate $g$ with a vocabulary $\tau$ in a circuit $C$ defines
% a labelled $\rho$-structure, where the labels are the syntactic-equivalence
% classes of the children of $g$.

% the restriction to transparency enables us to decide many natural circuit
% properties in polynomial-time (e.g.\ the orbit of a gate, whether the circuit
% is symmetric, etc.). important point is that in order to decide many
% properties of the circuit that we need in order to prove the main result
% (e.g.\ )

% In this section we justify this restriction, showing first that the a great
% number of very useful circuit properties (e.g. syntactic-equivalence) are
% polynomial-time decidable for transparent circuits and, second, that deciding
% most of these circuit properties for general circuits (with non-symmetric
% gates) is at least as hard as the graph-isomorphism problem. Since the proof
% of the main theorem makes use of the polynomial-time decidability of many of
% these properties, these results together justify the inclusion of the
% transparency restriction

\subsection{Tractable Properties of Transparent Circuits}
In the this subsection we show that many useful circuit properties are
polynomial-time decidable for transparent circuits. First, we prove that the
syntactic-equivalence relation is polynomial-time computable for transparent
circuits. We use this result to show that transparent circuits can be
transformed in polynomial-time into equivalent circuits with unique labels. We
show that for circuits with unique labels we can compute in polynomial-time the
action of a given automorphism on the circuit, the orbits and supporting
partitions of each gate, as well as the orbits and supporting partitions of each
element of the universe of each gate. We make use of all of these results in
order to define a translation from circuit families to formulas of $\FPR$ in
Section~\ref{sec:circuits-to-formulas}.

% The purpose of this section in then to show that we may transform tranparent
% circuits to circuits with unique labels, and since circuits with unique
% extension

% the function that computes circuits with unique labels

% We show that the set of transparent circuits, the set of circuits with unique
% labels, and the set of symmetric transparent circuits are all polynomial-time
% decidable. Moreover, we show that for circuits with unique labels

% and hence that the the action of a and the orbit and supporting partition of a
% gate are polynomial-time computable, . We also present polynomial-time
% algorithms for computing the action of a permutation on a symmetric circuit,
% as well as for computing the orbits and supports of gates, and elements of
% universes of gates, in a circuit.

% We use the polynomial-time decidability of many of these properties throughout
% this paper, and particularly for our translation of circuits into formulas.
% For example, the polynomial-time decidability of the syntactic-equivalence
% relation enables us to transform transparent circuits into circuits with
% unique labels (and so unique extensions) in polynomial-time. This result
% allows us, when working with $\PT$-uniform families of transparent circuits,
% to assume without a loss of generality that these circuits have unique labels
% and unique extensions. This is a crucial step in the proof of our main result,
% as the unique labels requirement is needed to apply the support theorem, and
% the unique extensions property is necessary for applying (and even defining) a
% great many technical tools (e.g. supports and supporting partitions) used
% extensively throughout this paper.

@@

\begin{lem}
  There is an algorithm that takes as input a transparent circuit $C$ and
  outputs the syntactic-equivalence relation on the gates of $C$. The algorithm
  runs in time polynomial in the size of $C$.
  \label{lem:transparent-syntactic-equiv}
\end{lem}
\begin{proof}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a transparent
  $(\mathbb{B}, \rho)$-circuit of order $n$. We define the syntactic-equivalence
  relation on $G$ by induction on the depth of a gate. We note that this
  definition by induction can be implemented as a construction, with the
  induction defining a dynamic program.

@@

  The above algorithm can be implemented by a dynamic program with the number of
  iterations bounded by the number of pairs of gates in $C$, i.e.\ ${\vert C
    \vert} \choose{2}$. It remains to show that there is an algorithm that runs
  in time polynomial in the size of the circuit and takes as input two gates $g$
  and $h$ in $C$ and the syntactic-equivalence relation for all gates of depth
  less than either $g$ or $h$ and returns whether or not $g \equiv h$. We now
  sketch a definition of this algorithm.

  Let $g$ and $h$ be two internal gates in $C$ and suppose we have defined the
  syntactic-equivalence relation for all gates of depth less than the depth of
  either $g$ or $h$. We check all of the conditions for syntactic-equivalence
  with the exception of the isomorphism condition, and any of them fail to hold
  we halt and output that $g \not\equiv h$.

  We now define a function $f: \ind(g) \rightarrow \ind(h)$ as follows. Let
  $\{x_1, \ldots, x_{\vert \ind(g)\vert}\}$ be an ordering of $\ind(g)$. For
  each $i \in \vert \ind(g) \vert$ select $y_i \in \ind(h)$ such that $y_i \neq
  y_j$ for all $j \in [i-1]$ and $L(g)(x_i) \equiv L(h)(y_i)$. If at any point
  we cannot select an appropriate $y_i$, halt and output that $g \not\equiv h$.
  For each $i \in \vert \ind(g) \vert$ let $f(x_i) = y_i$. From the definition
  of the function $f$ is injective. Check if $f$ is surjective. If not, halt and
  output that $g \not\equiv h$. We thus have that $f$ is a bijective function
  and for all $x \in \ind(g)$, $L(g)(x) \equiv L(h)(f(x))$. Check if both $g$
  and $h$ are both symmetric gates. If so, $f$ witnesses the required
  isomorphism-equivalence and so we halt and output that $g \equiv h$.

  We thus have that $g$ and $h$ are non-symmetric gates. Then, since $C$ is
  transparent, both $g$ and $h$ have unique labels. In that case the function
  $f$ is the only function such that for all $x \in \ind(g)$, $L(g)(x) \equiv
  L(h)(f(x))$. It follows that $g \equiv h$ if, and only if, $f$ defines an
  isomorphism from $\str{g}$ to $\str{h}$. This is easy to check.
\end{proof}

It is common approach to define a new structure by \emph{quotienting} a given
structure by some equivalence relation. We now define what it means to take a
quotient of a circuit by the syntactic-equivalence relation. Intuitively, we
think of a quotient of a given circuit as being formed by `merging' each
syntactic-equivalence class into a single gate and including a wire between two
gates in the quotient circuit if, and only if, there is a wire between an
element of each equivalence class.

@@

We should note that the there is no obvious quotienting operation that
associates with each circuit a unique quotient circuit. However, it is easy to
see that if $C$ and $C'$ are distinct quotients of a given circuit then the only
point where they differ is in the definition of their respective labelling
functions. But it is not hard to see that for every gate $g$ in $C$, $L(g)$ and
$L'(g)$ are isomorphism-equivalent. It follows that these two circuits are
structurally identical in a very strong sense. Indeed, while we have not
formally defined what it means for two circuits to be isomorphic, under any
reasonable definition we would have that any two quotients of a given circuit
are isomorphic.

We now show that taking the quotient of a circuit preserves important
properties, including the function computed by the circuit, the symmetry of the
circuit, and whether the circuit has unique labels. We also show that the
quotient of a circuit is reduced, and hence has unique children (and so, from
Proposition~\ref{prop:unique-children-unique-extensions}, unique extensions).

@@

  Let $g \in G$. Suppose $[g]$ has unique labels. There exists $h \in [g]$ such
  that $L_\equiv([g]) = L(h) /_\equiv$. Since $L_\equiv([g])$ is injective,
  $L(h) /_\equiv$ must be injective and so $L(h)$ must be injective and no two
  child gates of $L(h)$ must be syntactically-equivalent. It follows that $h$
  has unique labels. Since $h \equiv g$ and $h$ has unique labels, it follows
  that $g$ has unique labels. Suppose $g$ has unique labels. Let $h \in [g]$ be
  such that $L_\equiv([g]) = L(h) /_\equiv$. Since $h \equiv g$ and $g$ has
  unique labels $h$ has unique labels and so $L_\equiv([g])$ has injective
  labels. Since each equivalence class in $C_\equiv$ is a singlton it follows
  that $[g]$ has unique labels.
\end{proof}

In is not hard to show that there is a polynomial-time computable function that
maps a transparent circuit to a quotient of that circuit. To see this, recall
that from Lemma~\ref{lem:transparent-syntactic-equiv} we can compute the
syntactic-equivalence classes of a transparent circuit in polynomial-time. We
can thus define a quotient circuit by picking representatives from each
syntactic-equivalence class and then applying the definition of a quotient
circuit in the obvious manner.

We now show that transparent circuits can be transformed in polynomial time into
equivalent circuit with unique labels, and hence unique extensions, and that
this transformation preserves important properties.

\begin{lem}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a transparent
  $(\mathbb{B}, \rho)$-circuit. There is an algorithm that that takes in such a
  circuit and outputs a $(\mathbb{B} \cup \mathbb{B}_{\std}, \rho)$-circuit $C'$
  such that $C$ and $C'$ compute the same function, $C'$ has unique labels, and
  if $C$ is symmetric then $C'$ is symmetric. Moreover, this algorithm runs in
  time polynomial in the size of the input circuit.
  \label{lem:transparent-unique}
\end{lem}
\begin{proof}
  Let $C_0 := \langle G_0, \Omega_0, \Sigma_0, \Lambda_0, L_0 \rangle$ be the
  quotient of $C$. If $C_0$ does not contain the constant gates $g_0$ and $g_1$
  such that $\Sigma(g_0) = 0$ and $\Sigma (g_1) = 1$ we construct a new circuit
  from $C_0$ by adding gates. We abuse notation and also call this new circuit
  $C_0$. The addition of these constant gates to the circuit does not alter the
  function computed by the circuit, nor does it effect the symmetry of the
  circuit or weather it has unique labels.

@@

  We now define the circuit $C_1$. Let $C_1 := \langle G_1, \Omega_1, \Sigma_1 ,
  \Lambda_1, L_1 \rangle$ be defined as follows. Let $G_1 = G_0 \uplus \{
  g_\lor\}$. Let $\Omega_1 = \Omega_0$ and $\Lambda_1 = \Lambda_0$. Let $g \in
  G_1$. If $g = g_\lor$ then $\Sigma_1 (g) = \OR[2]$ with $L_1(g)(1) = g_0$ and
  $L_1(g)(2) = g_1$. If $g \in G_0$, $\Sigma(g) = \OR[2]$ and $H_g = \{g_0,
  g_1\}$, then $\Sigma_1 (g) = \OR[3]$, $L_1(g)(1) = L_0(g)(1)$, $L_1(g)(2) =
  L_0(g)(2)$, and $L_1(g)(3) = g_\lor$. If $g \in G_0$ and $g = \AND[k]$ for
  some $k \in \nats$ then $\Sigma_1(g) = \AND[k+1]$ and $L_1(g)(i) = L_0(g)(i)$
  for all $i \in [k]$ and $L_1(g)(k+1) = g_\lor$. Otherwise let $\Sigma_1(g) =
  \Sigma_0(g)$ and $L_1(g) = L_0(g)$.

  Stated more informally, we define $C_1$ from $C_0$ by adding in an $\OR$-gate
  $g_\lor$ that always evaluates to one, and then adding a wire from that gate
  to all $\AND$-gates in the circuit (and also a wire from $g_\lor$ to any
  similar gate that may already exist in the circuit in order to ensure that
  $g_\lor$ is part of a singleton syntactic-equivalence class in $C_1$). The
  important point to note is that each $\AND$-gate in $C_1$ has fan-in at least
  two. We will construct $C'$ from $C_1$ by adding in a number of $\AND$-gates
  with fan-in one. Therefore, since all of the $\AND$-gates in $C_1$ have fan-in
  two, it will follow that none of these new gates are syntactically-equivalent
  to any gate in $C_1$. We now show that $C_1$ and $C_0$ compute the same
  function and if $C_0$ is symmetric then $C_1$ is symmetric.

@@

  Suppose $C$ is symmetric. From Lemma~\ref{lem:quotient-circuits-preserve} it
  follows that $C_0$ is symmetric. Let $\sigma \in \sym_n$ and let $\pi_0 \in
  \aut(C_0)$ be an extension of $\sigma$. Let $\pi_1 : G_1 \rightarrow G_1$ such
  that $\pi_1 (g) = \pi_0(g)$ for all $g \in G_0$ and $\pi_1 (g_\lor) = g_\lor$.
  It is easy to see that $\pi_1$ is an automorphism of $C_1$ extending $\sigma$.
  It follows that $C_1$ is symmetric.

  We now show that $C_1$ is reduced. We have from
  Lemma~\ref{lem:quotient-circuits-preserve} that $C_0$ is reduced. Since
  $g_\lor$ is the only two input $\OR$-gate with exactly the two constant gates
  as children, $g_\lor$ is contained in a singleton syntactic-equivalence class.
  It can be shown by induction that if $g, g' \in G_0$ are syntactically
  equivalent in $C_1$ then they must be syntactically-equivalent in $C_0$. It
  follows from these two observations that if $g, g' \in G_1$ are
  syntactically-equivalent in $C_1$ then $g = g'$. We thus have that each
  syntactic-equivilence class in $C_1$ is a singleton, and so $C_1$ is reduced.
